\documentclass{article}

\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}

\title{CS536 Final Project}
\author{Author}
\date{}

\begin{document}
\maketitle

\section{Motivation}
Throughout this course, we have mostly considered learning settings where there is a single entity handling training and testing.
However, these techniques are not always applicable to scenarios like edge computing and IoT devices.
Often, there may be multiple devices with disjoint data sets that need to be aggregated to train a shared model.
These devices may also have bandwidth or privacy requirements, preventing users from directly aggregating data then training.
This is addressed by techniques in federated learning \cite{yang2019}.

There are several ways data can be split among multiple devices.
In the first case, each device can hold a subset of the data, recording different events with identical features.
This is the problem that \underline{horizontal federated learning} solves \cite{li2020}.
Alternatively, perhaps each device records some information about the same event, but they hold a disjoint set of features.
We address this problem with \underline{vertical federated learning} \cite{feng2020}.

\section{Vertical Federated Learning}
In a general learning problem, we have $\mathbf x\in\mathbb R^m$ as an input vector and we want to learn some function $f:\mathbb R^m\to\mathbb R$ and predict the value $y=f(\mathbf x)$.
Consider the $k$ party case, where we have $m=\sum_{i=1}^km_i$, and $\mathbf x=\bigoplus_{i=1}^m\mathbf x_i\in\bigoplus_{i=1}^m\mathbb R^{m_i}$.
Party $i$ can only see $\mathbf x_i,y$, and needs to cooperate with other parties to predict $f(\mathbf x_1\oplus\mathbf x_2)$ without divulging their own data.
We implement a simplified version of \cite{wan2007} for reasons we discuss later.

For simplicity, consider the case of linear regression.
Take $f(\mathbf x)=\mathbf x\cdot\mathbf w=\sum_{i=1}^k\mathbf x_i\cdot\mathbf w_i$ for $\mathbf w\in\mathbb R^m$, where $\mathbf w=\sum_{i=1}^k\mathbf w_i\in\bigoplus_{i=1}^m\mathbb R^{m_i}$ and party $i$ only sees $\mathbf w_i$.
Given some $\mathbf x_i$, $y$, party $i$ will compute and broadcast $\mathbf x_i\cdot\mathbf w_i$, then collect $\hat y=\sum_{i=1}^k\mathbf x_i\cdot\mathbf w_i$ from other parties.
The goal is to minimize the squared error $\mathcal L=(y-\hat y)^2$.

To train this model, some data $\mathbf X=(\mathbf x^1,\ldots,\mathbf x^N)^\intercal\in\mathbb R^{N\times m}$ and output $\mathbf y(y_1,\ldots,y_N)^\intercal\in\mathbb R^N$ is partitioned to $\mathbf X_i\in\mathbb R^{N\times m_1}$.
Party $i$ obtains $\mathbf X_i$, $\mathbf y$, and initializes $\mathbf w_i$ to a random vector.
Then we run several rounds of gradient descent, adjusting $w_l$ for each feature $l\in[m]$ as follows.
\begin{align}
    \frac{\partial\mathcal L}{\partial w_l}=&
    \frac{\partial}{\partial w_l}\sum_{i=1}^N\left(\sum_{j=1}^m \mathbf x_i(j)\cdot w_j-y_i\right)^2\\
    =&\sum_{i=1}^N
    2\mathbf x_i(l)\left(\sum_{j=1}^m \mathbf x_i(j)\cdot w_j-y_i\right)\\
    =&\sum_{i=1}^N2\mathbf x_i(l)(\mathbf x_i\cdot\mathbf w-y_i)
\end{align}

The key insight here is that the $\mathbf x_i\cdot\mathbf w-y_i$ can be computed without view of each individual feature, thus updating $w_l$ only requires $\mathbf x_i(l)$.
The party that needs to update $w_l$ also owns $\mathbf x_i(l)$, so federated gradient descent is possible.

\subsection{Gradient Inversion}
Upon closer inspection, the above method is functionally equivalent to gradient descent linear regression, but we claimed that it allowed data to be hidden from other parties.

\textbf{WIP}


\bibliographystyle{plain}
\bibliography{report.bib}

\end{document}